### 2013-11-30 paper/main.tex
engine model with no control working
With the torque input is set to zero a transfer function
representing the engine model can be constructed.
This should be suitable for building control systems
using pole placements or other methods.

	\node[gain] 	(gain1)	[]					{$1.699$};
	\node[sumnp]  	(sum1)	[right=of gain1] 	{};
	\node[gain] 	(gain2)	[right=of sum1]		{$\dfrac{8.5683}{z^2-0.9025z}$};
	\node[sumpn]  	(sum2)	[right=of gain2] 	{};
	\node[gain] 	(gain3)	[right=of sum2]		{$\dfrac{2.98}{z - 0.9534}$};
	\node[gain] 	(gain4)	[below=of sum2]		{$0.000093$};

@@

Typical control system models have a single input and a single output
\footnote{One input/output could also be a vector.}.
However this model is different in that it takes
two inputs: control and torque (Figure \ref{fig:lem}).
One way to reduce the model to a single input single output is to hold
one of the inputs constant.
However neither of the resulting use cases would make any sense to
a designer whose primary goal is to stabilize engine rpm.
The first case, with torque constant, would indicate the response for
control inputs.
But this is pointless because the goal of idle stabilization is to
regulate the rpm given changes in torque.
The second case, with control held constant, would indicate the response
for a change in torque.
But again this would provide no useful way to stabilize engine rpm.
Both inputs are needed in order to construct any viable control system.

\section{Regulator Control System}

@@

%And the open loop controller, with a fixed input duty cycle, is shown
%in Figure \ref{fig:openloop}.
%Notice that in this example the torque has transient characteristics and
%varies over time.
%The output of this system is show in Figure \ref{fig:olplot}.

@@

\begin{align}
	y &= u + r \notag \\
	r &= y \cdot z^{-1} \notag \\
	y &= u + y z^{-1} \notag \\
	y &= y(1 - z^{-1}) \notag \\
	\Aboxed{ \frac{y}{u} &= \frac{1}{1 - z^{-1}} } \label{eq:dc2s}
\end{align}

@@

@@

\clearpage
\section{Linear Engine Model Plot Matlab Script}
\label{app:olplot}
\lstinputlisting[language=Matlab]{../matlab/linear_engine_model_plot.m}

\section{Linear Engine Model Initialization Matlab Script}
\lstinputlisting[language=Matlab]{../matlab/linear_engine_model_init.m}

>>>>

\begin{tikzpicture}[>=triangle 60, scale=0.7,
	node distance=14mm, auto]

	\node[gain] 	(gain1)	[]					{$1.699$};
	\node[sumnp]  	(sum1)	[right=of gain1] 	{};
	\node[gain] 	(gain2)	[right=of sum1]		{$\dfrac{8.5683}{z^2-0.9025z}$};
	\node[sumpn]  	(sum2)	[right=of gain2] 	{};
	\node[gain] 	(gain3)	[right=of sum2]		{$\dfrac{2.98}{z - 0.9354}$};
	\node[gain] 	(gain4)	[below=of sum2]		{$0.00093$};

@@

In order to build a controller the model needs to be simplified
but the presence of two inputs complicates matters.
To resolve this issue the torque can be set zero.
With this change it can be simplified by first recognizing that
it matches the form shown in Figure \ref{fig:dd} which
has the transfer function in Equation \ref{eq:dd}.

% {{{ dd figure
\begin{figure}[h!]

\begin{center}
\begin{tikzpicture}[>=triangle 60, scale=0.8,
	node distance=20mm, auto]

	\node[gain]		(K) 	[] 				{$K$};
	\node[sumnp]	(sum1) 	[right of=K] 	{};
	\node[gain]		(Gz) 	[right of=sum1]	{$G(z)$};
	\node[gain]		(Dz) 	[below of=Gz] 	{$D(z)$};

	\draw [<-] (K) -- ++(-20mm,0) node[above]{$u$};
	\draw [->] (K) -- (sum1);
	\draw [->] (sum1) -- (Gz);
	\draw [<-] (sum1) |- (Dz);
	\draw [->] (Gz) -- ++(35mm,0) node[above]{$y$};
	\draw [->] (Gz) -- ++(20mm,0) |- (Dz.east);
\end{tikzpicture}
\end{center}

\caption{Direct Design system with $K$ as a scaling input, $G$ is
the plant and $D$ is the controller.}
\label{fig:dd}
\end{figure}
% }}}

\begin{align}
\frac{y}{u} &= K \frac{B \alpha}{A \alpha + B \beta} \label{eq:dd}
\end{align}

First taking the parts from the engine model (Figure \ref{fig:lem}).

\begin{align*}
	\frac{B}{A} &= \frac{8.5683}{(z^2 - 0.9025z)} \frac{2.98}{(z - 0.9354)} \\
	K &= 1.699 \\
	\frac{\beta}{\alpha} &= 0.00093 
\end{align*}

Then substituting them in to Equation \ref{eq:dd} results in
the simplified from of Equation \ref{eq:lems}.

\begin{align}
	\frac{y}{u} &= \frac{(1.699) (8.5683) (2.98) (1)}
		{(z^2 - 0.9025z)(z - 0.9354)(1) + (8.5683)(2.98)(0.00093)} \notag \\
	\Aboxed{ \frac{y}{u} &= \frac{43.38}{z^3 - 1.838 z^2 + 0.8442 z + 0.02375}}
		\label{eq:lems}
\end{align}

\begin{samepage}
It is still necessary to address the delta inputs and outputs.
It can be seen in Figure \ref{fig:nodelta} there is a beneficial
side effect that makes this trivial.  The transform that converts
from steady state to a delta cancels
with the transform that converts from a delta to steady state
\footnote{If the torque input was non zero the transforms would not
cancel.}.  Therefore the final transform is still Equation \ref{eq:lems}.
An example of the output response with no control is shown in
Figure \ref{fig:em_plot}.
\end{samepage}

% {{{ fig:nodelta
\begin{figure}[!htbp]

\begin{center}
\begin{tikzpicture}[>=triangle 60,
	node distance=30mm, auto]

\node[gain] (gain1) [] {$1 - z^{-1}$};
\node[gain] (engine) [right of=gain1] {engine};
\node[gain] (gain3) [right of=engine] {$\dfrac{1}{1 - z^{-1}}$};

\draw [->] (-25mm,0) node[above]{$u$} -- (gain1.west);
\draw [->] (gain1.east) node[above,xshift=6mm]{$\Delta u$} -- (engine);
\draw [->] (engine.east) node[above,xshift=6mm]{$\Delta y$} -- (gain3);
\draw [->] (gain3.east) -- ++(20mm,0) node[above]{$y$};
\end{tikzpicture}
\end{center}

\caption{Engine model with transforms for converting from steady state
to delta and vice versa.}
\label{fig:nodelta}

\begin{figure}[htbp!]
\begin{center}
\includegraphics[scale=0.7]{../octave/engine_model_plot}
\end{center}
\caption{Output response of engine model with no control.
The control can range from zero to one.}
Here the control is constant at 0.65 which resulted in a stable
rpm near 900.
The Matlab script is provided in Appendix \ref{app:em_matlab}.
\label{fig:em_plot}
\end{figure}

% see octave/engine_model.m for these calculations

\clearpage
\section{Regulator Control System}

@@

@@

\begin{align}
	y &= u + r \notag \\
	r &= y \cdot z^{-1} \notag \\
	y &= u + y z^{-1} \notag \\
	u &= y(1 - z^{-1}) \notag \\
	\Aboxed{ \frac{y}{u} &= \frac{1}{1 - z^{-1}} } \label{eq:dc2s}
\end{align}

@@

\begin{tikzpicture}[>=triangle 60]
	\node[gain] (gain1) [] {$\dfrac{1}{1 - z^{-1}}$};
\draw [->] (-20mm,0) node[above,yshift=1mm]{$u$} -- (gain1.west);
\draw [->] (gain1.east) -- (20mm,0) node[above,yshift=1mm]{$y$};
\end{tikzpicture}

@@

\clearpage
\section{Engine Model Matlab Source}
\label{app:em_matlab}

\lstinputlisting[
	caption={Matlab source to calculate the engine model system with
		zero torque.}
	]{../octave/engine_model.m}

\lstinputlisting[
	caption={Matlab source to plot the response of engine model
		with no control.}
	]{../octave/engine_model_plot.m}
